Massively Parallel Monte Carlo Methods for Discrete Linear and Nonlinear Systems

### 2012-11-01 doc/prelim/presentation/prelim_presentation.tex
added conclusion
spell check on presentation

  \begin{itemize}
  \item Widely used in practice
  \item Krylov methods require only the action of the operator
  \item Parallelism achieved through a handful of operations
  \item Global reduction operations observed not to impede scalability
    \begin{itemize}
    \item Dot product
    \item Vector norms
    \end{itemize}
  \item Nearest neighbor computations have poor algorithmic strong
    scaling
    \begin{itemize}
    \item Matrix-vector multiply
    \item Weak scaling is better
    \end{itemize}
  \item Short and long reccurrence relations available for
    orthogonalization
  \end{itemize}

@@

  \begin{itemize}
  \item Expand the Nuemann series
  \end{itemize}

@@

  \begin{itemize}
  \item Published work to date has used a physics-based formulation
    \begin{itemize}
    \item Radiation transport equations used as model system
    \item Transition probabilities built from problem-specific
      parameters
    \item Probabilities and weights must be rederived for each new
      equation set
    \end{itemize}
  \item Desire a generalization for all linear operator equations
    \begin{itemize}
    \item Requires a general parallel framework
    \item Requires implementation with a general linear algebra
      framework
    \item Operator, vector, and graph abstractions
    \end{itemize}
  \item Neumann-Ulam solvers and MCSA implemented using the Trilinos
    Petra frameworks
  \item Can be leveraged in modern physics implementations
  \end{itemize}

@@

  \begin{itemize}
  \item Sigel observed $\lambda$ to be bound empirically to the mean
    free path of the system
  \item Optimum ratio of local to global domain size exists
  \item Experiments show feasibility for large-scale calculations
    \begin{itemize}
    \item Not bandwidth or latency limited in load-balanced case
    \end{itemize}
    \pause
  \item How do we define $\lambda$ empirically for linear operator
    equations? 
  \item Full-clip roullette for the MCSA correction?
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Reproducability}

@@

  \begin{itemize}
  \item 2005 work by Gentile outlines reproducability in parallel
    Monte carlo
  \item Maintaining random number sequences
  \item Floating point addition does not commute
  \item Useful for benchmarking and unit testing
  \end{itemize}

@@

Questions
  \begin{itemize}
  \item How much domain overlap is suitable for linear operator
    equations? 
  \item Is memory a limitation for overlap and replication?
  \item Do the linear operator properties help select overlap?
  \item Will full-clip roullette perturb the MCSA solution?
  \item How much does MSOD facilitate scaling?
  \end{itemize}

@@

  \begin{itemize}
  \item Modern FAD packages are parallelized using element-level
    assembly
  \item This work will generate a parallel MCSA solver
  \item Application of the Newton correction requires a parallel
    vector update
  \item Convergence checks through parallel norm comptutation
  \end{itemize}

@@

  Parallel Performance and Numerical Accuracy Studies for the MCSA
  Method
  \begin{itemize}
  \item Characterize Neumann-Ulam parameters required for good
    correction
  \item Feasibility of MCSA with full-clip Neumann-Ulam
  \item Performance for assymmetric systems
  \item Memory usage compared to Krylov methods
  \end{itemize}

@@

>>>>

  \begin{itemize}
  \item Widely used in practice
  \item Krylov methods require only the action of the operator
  \item Parallelism achieved through a handful of operations
  \item Global reduction operations observed not to impede scalability
    \begin{itemize}
    \item Dot product
    \item Vector norms
    \end{itemize}
  \item Nearest neighbor computations have poor algorithmic strong
    scaling
    \begin{itemize}
    \item Matrix-vector multiply
    \item Weak scaling is better
    \end{itemize}
  \item Short and long recurrence relations available for
    orthogonalization
  \end{itemize}

@@

  \begin{itemize}
  \item Expand the Neumann series
  \end{itemize}

@@

  \begin{itemize}
  \item Published work to date has used a physics-based formulation
    \begin{itemize}
    \item Radiation transport equations used as model system
    \item Transition probabilities built from problem-specific
      parameters
    \item Probabilities and weights must be re-derived for each new
      equation set
    \end{itemize}
  \item Desire a generalization for all linear operator equations
    \begin{itemize}
    \item Requires a general parallel framework
    \item Requires implementation with a general linear algebra
      framework
    \item Operator, vector, and graph abstractions
    \end{itemize}
  \item Neumann-Ulam solvers and MCSA implemented using the Trilinos
    Petra frameworks
  \item Can be leveraged in modern physics implementations
  \end{itemize}

@@

  \begin{itemize}
  \item Siegel observed $\lambda$ to be bound empirically to the mean
    free path of the system
  \item Optimum ratio of local to global domain size exists
  \item Experiments show feasibility for large-scale calculations
    \begin{itemize}
    \item Not bandwidth or latency limited in load-balanced case
    \end{itemize}
    \pause
  \item How do we define $\lambda$ empirically for linear operator
    equations? 
    \pause
  \item Full-clip roulette for the MCSA correction?
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Reproducibility}

@@

  \begin{itemize}
  \item 2005 work by Gentile outlines reproducibility in parallel
    Monte Carlo
  \item Maintaining random number sequences
  \item Floating point addition does not commute
  \item Useful for benchmarking and unit testing
  \end{itemize}

@@

Questions
  \begin{itemize}
  \item How much domain overlap is suitable for linear operator
    equations? 
  \item Is memory a limitation for overlap and replication?
  \item Do the linear operator properties help select overlap?
  \item Will full-clip roulette perturb the MCSA solution?
  \item How much does MSOD facilitate scaling?
  \end{itemize}

@@

  \begin{itemize}
  \item Modern FAD packages are parallelized using element-level
    assembly
  \item This work will generate a parallel MCSA solver
  \item Application of the Newton correction requires a parallel
    vector update
  \item Convergence checks through parallel norm computation
  \end{itemize}

@@

  Parallel Performance and Numerical Accuracy Studies for the MCSA
  Method
  \begin{itemize}
  \item Characterize Neumann-Ulam parameters required for good
    correction
  \item Feasibility of MCSA with full-clip Neumann-Ulam
  \item Performance for asymmetric systems
  \item Memory usage compared to Krylov methods
  \end{itemize}

@@

  \begin{itemize}
  \item Proposed research and development of new Monte Carlo methods
    \begin{itemize}
    \item Parallelization of Monte Carlo methods for linear systems
    \item FANM for nonlinear systems
    \end{itemize}
  \item Verification through benchmarks
  \item Numerical experiments for understanding
  \item Challenge problem for application
  \item Directed towards hard problems in nuclear reactor analysis:
    \begin{itemize}
    \item Application to multiphysics
    \item Potential improvements for scalability
    \item Potential improvements for memory consumption
    \item Looking forward to exascale
    \end{itemize}
  \end{itemize}
